---
title: "R Notebook"
output: html_notebook
---

## 1. Data Preparation

```{r}

library("xts")
# install.packages("dplyr")
library('dplyr') # data manipulation
# library('readr') # input/output
# library('data.table') # data manipulation
# install.packages('tibble')
library('tibble') # data wrangling
# library('tidyr') # data wrangling
# library('stringr') # string manipulation
# library('forcats') # factor manipulation


# install.packages('tidytext')
library('tidytext')
# install.packages("magrittr")
library('magrittr')
# install.packages("tidyr")
library("tidyr")
# install.packages("ggplot2")
library("ggplot2")

# install.packages("plotly")
library("plotly")

# install.packages("pamr")
library("pamr")

library(wordcloud)

source("../lib/multiplot.R")

# install.packages("gridExtra")
library("gridExtra")
```



```{r}

# We want to treat each column as characters, not factors, except for the author column.
spooky <- read.csv('../data/spooky.csv', colClasses = 'character')
spooky <- as.tibble(spooky)
spooky$author <- as.factor(spooky$author)
summary(spooky)

```

## 2. Pronoun Occurence 
```{r fig.keep="last"}
# First use tidytext function to drop the punctuations and tokenize our file.
# We use bigrams because we want the word following the pronouns "he" and "she".
pronouns <- c("he","she")
data <- spooky %>%
  unnest_tokens(word, text, token = "ngrams", n = 2, to_lower = TRUE, drop = TRUE)


data_counts <- data %>%
  count(word, sort = TRUE) %>%
  separate(word, c("word1","word2"), sep = " ", remove = TRUE) %>%
  filter(word1 %in% pronouns) %>%
  filter(word2 != "he") %>%
  count(word1, word2, wt = n, sort = TRUE) %>%
  rename(counts = "nn")

all_po <- data.frame(
  count(data_counts, word1)
)
all_po <- all_po %>%
  rename(counts = "n",
         pronouns = "word1")

png("../figs/po1.png")
ggplot(all_po, aes(x="", y=counts, fill=pronouns))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0) +
  ggtitle("pronoun occurence in total")

dev.off()

data_counts_by_author <- data %>%
  count(author, word, sort = TRUE) %>%
  separate(word, c("word1","word2"), sep = " ", remove = TRUE) %>%
  filter(word1 %in% pronouns) %>%
  filter(word2 != "he") %>%
  rename(counts = "n")

MWS_po <- data_counts_by_author %>%
  filter(author == "MWS") %>%
  count(word1) %>%
  rename(pronoun = "word1", counts = "n")
HPL_po <- data_counts_by_author %>%
  filter(author == "HPL") %>%
  count(word1) %>%
  rename(pronoun = "word1", counts = "n")
EAP_po <- data_counts_by_author %>%
  filter(author == "EAP") %>%
  count(word1) %>%
  rename(pronoun = "word1", counts = "n")

p1 <- ggplot(MWS_po, aes(x="", y=counts, fill=pronoun))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0) +
  ggtitle("MWS pronoun occurence")
p2 <- ggplot(HPL_po, aes(x="", y=counts, fill=pronoun))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0) +
  ggtitle("HPL pronoun occurence")
p3 <- ggplot(EAP_po, aes(x="", y=counts, fill=pronoun))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0) +
  ggtitle("EAP pronoun occurence")

layout <- matrix(c(1,2,3), 3,1, byrow = TRUE)
png("../figs/po2.png")
multiplot(p1, p2, p3, layout = layout)
dev.off()

```


## 3. Gender Actions
```{r}

word_ratios <- data_counts %>%
    group_by(word2) %>%
    filter(sum(counts) > 10) %>%
    ungroup() %>%
    spread(word1, counts, fill = 0) %>%
    mutate_if(is.numeric, funs((. + 1) / sum(. + 1))) %>%
    mutate(logratio = log2(she / he)) %>%
    arrange(desc(logratio))   

# Equally likely:
equal <- word_ratios %>%
  mutate(abslogratio = abs(logratio)) %>%
  arrange(abslogratio)
words <- equal$word2
freqs <- equal$abslogratio

wordcloud(words, freqs, max.words = 14, min.freq = -0.25, scale = c(4,.5), vfont = c("sans serif","plain"), color = c("purple4", "red4", "black"))

# Large difference:

png("../figs/verbs1.png")
word_ratios %>%
    mutate(abslogratio = abs(logratio)) %>%
    group_by(logratio < 0) %>%
    top_n(15, abslogratio) %>%
    ungroup() %>%
    mutate(word = reorder(word2, logratio)) %>%
    ggplot(aes(word, logratio, color = logratio < 0)) +
    geom_segment(aes(x = word, xend = word,
                     y = 0, yend = logratio), 
                 size = 1.1, alpha = 0.6) +
    geom_point(size = 3.5) +
    coord_flip() +
    labs(x = NULL, 
         y = "Relative appearance after 'she' compared to 'he'",
         title = "Words paired with 'he' and 'she'",
         subtitle = "Women throw, sleep, and turn while men####") +
    scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
    scale_y_continuous(breaks = seq(-3, 3),
                       labels = c("0.125x", "0.25x", "0.5x", 
                                  "Same", "2x", "4x", "8x"))
dev.off()

# Separately:
# EAP

EAP_word_ratios <- data_counts_by_author %>%
  filter(author == "EAP") %>%
  group_by(word2) %>%
  filter(sum(counts) > 5) %>%
  ungroup() %>%
  spread(word1, counts, fill = 0) %>%
  mutate_if(is.numeric, funs((. +1)/ sum(. +1))) %>%
  mutate(logratio = log2(she / he)) %>%
  arrange(desc(logratio)) %>%
  mutate(abslogratio = abs(logratio)) %>%
  group_by(logratio < 0) %>%
  top_n(15,abslogratio) %>%
  ungroup() %>%
  mutate(word = reorder(word2, logratio))

jpeg("../figs/EAPverbs.jpeg", quality = 100)
ggplot(EAP_word_ratios, aes(word, logratio, color = logratio < 0)) +
  geom_segment(aes(x=word, xend = word,
                   y = 0, yend = logratio),
               size = 1.1, alpha = 0.6) +
  geom_point(size = 3.5) +
  coord_flip() +
  labs(x = NULL,
       y = "Relative appearance after 'she' compared to 'he' in EAP's novels",
       title = "Words paired with 'he' and 'she'") +
  scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
  scale_y_continuous(breaks = seq(-3, 3),
                     labels = c("0.125x", "0.25x", "0.5x", 
                                "Same", "2x", "4x", "8x"))
dev.off()

# HPL

HPL_word_ratios <- data_counts_by_author %>%
  filter(author == "HPL") %>%
  group_by(word2) %>%
  filter(sum(counts) > 5) %>%
  ungroup() %>%
  spread(word1, counts, fill = 0) %>%
  mutate_if(is.numeric, funs((. +1)/ sum(. +1))) %>%
  mutate(logratio = log2(she / he)) %>%
  arrange(desc(logratio)) %>%
  mutate(abslogratio = abs(logratio)) %>%
  group_by(logratio < 0) %>%
  top_n(15,abslogratio) %>%
  ungroup() %>%
  mutate(word = reorder(word2, logratio))

jpeg("../figs/HPLverbs.jpeg", quality = 100)
ggplot(HPL_word_ratios, aes(word, logratio, color = logratio < 0)) +
  geom_segment(aes(x=word, xend = word,
                   y = 0, yend = logratio),
               size = 1.1, alpha = 0.6) +
  geom_point(size = 3.5) +
  coord_flip() +
  labs(x = NULL,
       y = "Relative appearance after 'she' compared to 'he' in HPL's novels",
       title = "Words paired with 'he' and 'she'") +
  scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
  scale_y_continuous(breaks = seq(-3, 3),
                     labels = c("0.125x", "0.25x", "0.5x", 
                                "Same", "2x", "4x", "8x"))
dev.off()
 
# MWS

MWS_word_ratios <- data_counts_by_author %>%
  filter(author == "MWS") %>%
  group_by(word2) %>%
  filter(sum(counts) > 5) %>%
  ungroup() %>%
  spread(word1, counts, fill = 0) %>%
  mutate_if(is.numeric, funs((. +1)/ sum(. +1))) %>%
  mutate(logratio = log2(she / he)) %>%
  arrange(desc(logratio)) %>%
  mutate(abslogratio = abs(logratio)) %>%
  group_by(logratio < 0) %>%
  top_n(15,abslogratio) %>%
  ungroup() %>%
  mutate(word = reorder(word2, logratio))

jpeg("../figs/MWSverbs.jpeg", quality = 100)
ggplot(MWS_word_ratios, aes(word, logratio, color = logratio < 0)) +
  geom_segment(aes(x=word, xend = word,
                   y = 0, yend = logratio),
               size = 1.1, alpha = 0.6) +
  geom_point(size = 3.5) +
  coord_flip() +
  labs(x = NULL,
       y = "Relative appearance after 'she' compared to 'he' in MWS's novels",
       title = "Words paired with 'he' and 'she'") +
  scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
  scale_y_continuous(breaks = seq(-3, 3),
                     labels = c("0.125x", "0.25x", "0.5x", 
                                "Same", "2x", "4x", "8x"))
dev.off()

```




# Future directions for research
data cleaning:
1. Look at type of sentence, punctuations
2. n-gram instead of words; groups of words that represent a theme
3. length of sentence in words; types of words like nouns, adjectives, etc.
4. stemming/tization
5. any information in the stop words; 
6. studying the first and last words in the sentences
7. HPL makes up a lot of words
8. can we find specifically "british", "french" words for ms
9. Sentence generation

sentiment analysis:
1. sentiment analysis at sentence level instead of word level. What proportion of the sentence is positive?
2. k-means clusters on emotion, counts, or sentence. 
3. Are the results from the other lexicons different?
4. We're looking at the word counts for each emotion... is there some sort of normalization important here? eg. number of negative words in single sentence
5. What are the top few emotions represented by each author and are they the same across the authors?
6. Psycho profiles of the authors.

topic modeling:
1. visualizing author topics
2. determining the number of topics
3. topic model it at the corpus level, instead of sentence level.
4. other algorithms other than LDA
5. 


**clustering analysis



```{r}
gender <- data_counts_by_author %>%
  spread(word1, counts, fill = 0)
temp_male <- gender %>%
  spread(word2, he, fill = 0) %>%
  group_by(author) %>%
  summarise_if(is.numeric,sum) %>%
  ungroup() %>%
  rename(gender = "she") %>%
  mutate(gender = "male")
temp_female <- gender %>%
  spread(word2, she, fill = 0) %>%
  group_by(author) %>%
  summarise_if(is.numeric,sum) %>%
  ungroup() %>%
  rename(gender = "he") %>%
  mutate(gender = "female")
gender <- rbind(temp_male, temp_female)
gender$author <- as.character(gender$author)
first_row <- c("","",rep("word", 1134) %>%
  paste0(1:1134))
second_row <- c("", "", colnames(gender)[-c(1,2)])
gender <- rbind(first_row,second_row, gender) %>%
  as.matrix()
colnames(gender) <- NULL
gender <- t(gender)

write.table(gender, "../data/pam.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
gender.data <- pamr.from.excel("../data/pam.txt", ncols = 8, sample.labels = TRUE)
temp <- tableGrob(head(gender))
grid.newpage()
png("../figs/data.png")
grid.draw(temp)
dev.off()


# khan.data <- pamr.from.excel("../data/khan.txt", ncols = 65, sample.labels = TRUE)
# khan.train <- pamr.train(khan.data)
# khan.results <- pamr.cv(khan.train, khan.data)
# pamr.plotcv(khan.results)
# pamr.confusion(khan.results, threshold=4)
# pamr.plotcvprob(khan.results, khan.data, threshold=4.0)
# pamr.plotcen(khan.train, khan.data, threshold=4.0)
# pamr.geneplot(khan.train, khan.data, threshold=5.3)






gender.train <- pamr.train(gender.data)
gender.results <- pamr.cv(gender.train, gender.data)

## Plot the cross-validated error curves
png("../figs/graph1.png")
pamr.plotcv(gender.results)

## Compute the confusion matrix for a particular model (threshold=7.0) 
pamr.confusion(gender.results, threshold=7.0)

## Plot the cross-validated class probabilities by class
pamr.plotcvprob(gender.results, gender.data, threshold=7.0)

## Plot the class centroids
# words that matters
pamr.plotcen(gender.train, gender.data, threshold=2.0)

## Make a gene plot of the most significant genes
pamr.geneplot(gender.train, gender.data, threshold=3)



```


